---
title: "Biomass plots"
author: "Joe Glass-Katz"
date: "7/6/2018"
output: html_document
---
This file is to plot and explore the allometric equations taken from the Forest 2017 paper. The Pdf is located on the google drive
https://drive.google.com/drive/folders/1kODqAvB6VIYmyG0XjEuW1cKZjdAzC1iM?ogsrc=32
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(data.table)
```

make some plots
```{r}
#These are the averaged functions
pineAGB <- function(d,h) 0.0291*d^1.74165*h^1.16614 + 0.02029*d^1.33299*h^0.92887 + 0.02508*d^1.83773*h^0.54626 + 0.05227*d^1.28231*h^0.43275 
oakAGB <- function(d,h){0.11618*d^1.77395*h^0.68708 + 0.00827*d^2.54589 + 0.0502*d^1.97638*h^0.34229 + 0.08234*d^1.59396}

#Paramatrize using measured values
pinePlot <- ggplot(data = data.frame(x=0), aes(x=x)) + 
  stat_function(fun = pineAGB, args =list(h=10), aes(color = 'H=10')) +
  stat_function(fun = pineAGB, args =list(h=20), aes(color = 'H=20'))+
  stat_function(fun = pineAGB, args =list(h=30), aes(color = 'H=30'))+ scale_x_continuous(name = "Diameter (cm)", breaks = seq(5,60,10), limits = c(5,60))+
scale_y_continuous(name = "Estimated AGB")+
scale_color_manual(name = "Heights", values = c("blue", "green","deeppink"))+ ggtitle("AGB for Pine Trees")  

oakPlot <- ggplot(data = data.frame(x=0), aes(x=x)) + 
  stat_function(fun = oakAGB, args =list(h=5), aes(color = 'H=5')) +
  stat_function(fun = oakAGB, args =list(h=15), aes(color = 'H=15'))+
  stat_function(fun = oakAGB, args =list(h=25), aes(color = 'H=25'))+ scale_x_continuous(name = "Diameter (cm)", breaks = seq(5,60,10), limits = c(5,60))+
scale_y_continuous(name = "Estimated AGB")+
scale_color_manual(name = "Heights", values = c("blue", "green","deeppink"))+ ggtitle("AGB for Oak Trees")

#Now do foliage for pine and oak
pineFAGB <-function(d,h) 0.05227*d^1.28231*h^0.43275
oakFAGB <-function(d) 0.08234*d^1.59396

#Does the crown area scale at the same rate as the FAGB??

#Diameter to height equations (cm->m)
angiosperm <- function(d) 2.37*d^.53
gymnosperm <- function(d) 3.21*d^.6
Evergreen <- function(d) 3.25*d^.6
Deciduous <- function(d) 2.35*d^.53  


#DBH to height for Pine trees (RC25)
#file:///Users/jsg1/Desktop/Groundwater%20Research/RC25.pdf
#Model equation H =4.5 + exp(a0 + a1DBH^a2)
#Divide by 2.54 to convert to cm and multiply by .3048 tp get m
dougfirH <-function(D) (4.5 + exp(7.262195456 -5.899759104*(D/2.54)^ -0.287207389)) *.3048

ralderH <-function(D) (4.5 + exp( 4.41820972 -12.00274935*(D/2.54)^ -2.13835482))*.3048 

sspruH <-function(D) (4.5+exp( 5.404491308-6.570862442*(D/2.54)^ -0.819705048)) *.3048

hemlockH<-function(D) (4.5 + exp( 6.555344622-5.137174162*(D/2.54)^-0.3645508))*.3048 

rcedarH<-function(D) (4.5+exp( 7.232880669-5.746899904*(D/2.54)^-0.271564741))*.3048

wpineH<-function(D) (4.5 + exp( 7.946192109 -6.278973035*(D/2.54)^ -0.208892429))*.3048

pineMeanH <-function(x) (dougfirH(x) + hemlockH(x) + sspruH(x)+ rcedarH(x) +wpineH(x))/5 

PineHplot <- ggplot(data = data.frame(x=0), aes(x=x)) + 
  stat_function(fun = dougfirH, aes(color = "Dougfir")) +
  stat_function(fun = ralderH, aes(color = "Alder")) + 
  stat_function(fun = sspruH, aes(color = "Spruce")) +
  stat_function(fun = hemlockH, aes(color = "Hemlock")) +
  stat_function(fun = rcedarH, aes(color = "Cedar")) + 
  stat_function(fun = wpineH, aes(color = "WhitePine"))+
  stat_function(fun = pineMeanH, aes(color = "Average")) + 
  scale_x_continuous(name = "Diameter (cm)", breaks = seq(5,60,10), limits = c(5,60))+
  scale_y_continuous(name = "Estimated Height (m)")+
  scale_color_discrete(name = "Species" )+
  ggtitle("DBH to Height Gymnosperms")

#Generate points, flip the equation, regress to get an estimate
d <- seq(5,100,.01)
h <- pineMeanH(d)
df <- data.frame(d,h)
lmH<-lm(log(d)~ h, data =df)
plot(lmH$residuals)
#Simple exponential model for DBH pine is d = exp(1.938 + .05272h)
loglogH <-lm(log(d)~log(h), data=df)
plot(loglogH$residuals)
#Loglog is d = e^-0.769598 * h^1.320069
#more complicated using the equation from the paper (with start at where?)
#lower bound needs to be determined (change the 4.5)
Hfun<-function(a0,a1,a2,d){4.5 + exp(a0 + a1*d^a2)}
complexfit <- nls(h~Hfun(a0,a1,a2,d), start = list(a0=6,a1=-6,a2=-.3), data = df)
complexfit
a0<- 4.8565
a1 <- -10.2234
a2 <- -.5086
cfit <- function(d) 4.5 + exp(a0 + a1*d^a2)
genH <- function(d) exp(.769598)*d^.7575
lemH <- function(d) (log(d)-1.938) *(1/.05272)
PineHplot + 
  stat_function(fun = genH, aes(color ="loglog")) + 
  stat_function(fun = lemH, aes(color = "Log D") ) + 
  stat_function(fun = cfit, aes(color = "fitted model"))
#Fitted model clearly dominates. Use the complexfit for model

#make this "githubbable"


#Now let's do oak DBH to H
#There is a severe lack of data containing this information. Need to comb through the files in the google drive to see if there is something more usable there

```

#Relevant wood density & specific gravities
```{r}

pinusWD <- .433
quercusWD <- .65
#https://www.fs.fed.us/nrs/pubs/jrnl/2014/nrs_2014_chojnacky_001.pdf
pinusWSG <- .43
oakWSG <- .58

#Allometric equations from the above paper
#They take a generalized measurement
pinusAGB1<- function(D) exp(-2.6177+ 2.4638*log(D))
pinusAGB2 <-function(D) exp(-3.0506+ 2.6465*log(D))
pinusAGBa <-function(D) (exp(-2.6177+ 2.4638*log(D)) + exp(-3.0506+ 2.6465*log(D)))/2
quercusAGB1 <-function(D) exp(-2.1505 + 2.4410*log(D))

#an alternative oak AGB from https://www.sciencedirect.com/science/article/pii/S0961953413004935
quercusAGB2 <-function(d)  0.0683*d^2.5697
```

#Compare the two sets of allometric equations
```{r}
#Subbing in the height to diameter equations to solve for D let's compare these two sets of equations.
#First for pine
d <- seq(5,100,.01)
agb <- pinusAGBa(d)
df1 <- data.frame(d,agb)
agbfun <- function(d,b0,b1) exp(b0 + b1*log(d))
agpine <- nls(agb ~agbfun(b0,b1,d), start = list(b0 = -2.7, b1 = 2.5), data =df1)
agpine
b0 <- -2.886
b1 <- 2.571
#Put the first pine function just in terms of D
pine.No.H <- function(d) pineAGB(d=d,h=Hfun(a0,a1,a2,d))
#Simplify agbfun
agbfun1 <- function(d) exp(-2.866 +2.571*log(d))

#Compare these two functions
pineAGBCompare <- ggplot(data = data.frame(x=0), aes(x=x)) +
  stat_function(fun = pine.No.H, aes(color ="Formula1")) + 
  stat_function(fun = agbfun1, aes(color ="Formula2")) +
  scale_x_continuous(name = "Diameter (cm)", breaks = seq(5,60,10), limits = c(5,60))+
  scale_y_continuous(name = "Estimated AGB")+
  scale_color_discrete(name = "Model" )+
  ggtitle("DBH to AGB Gymnosperms(pines)")

pineAGBCompare
#We can take the average of these two equations to find a good neutral estimate for pine tree allometry. We just need to determine our ideal functional form and the rest is trivial.
#Straight power law is fine
pAGB.ave <- function(b0,b1,d) 

#Finding now the oak height to density equation
#for lack of a good oak average DBH to H we will use as a stand in the generalized angiosperm equation
oak.No.H <- function(d) oakAGB(d,angiosperm(d))
oakAGBCompare <- ggplot(data = data.frame(x=0), aes(x=x)) +
  stat_function(fun = oak.No.H, aes(color ="Formula1")) + 
  stat_function(fun = quercusAGB1, aes(color ="Formula2")) +
  stat_function(fun = quercusAGB2, aes(color ="Formula3")) +
  scale_x_continuous(name = "Diameter (cm)", breaks = seq(5,60,10), limits = c(5,60))+
  scale_y_continuous(name = "Estimated AGB")+
  scale_color_discrete(name = "Model" )+
  ggtitle("DBH to AGB Oaks")

oakAGBCompare
#While formula1 and formula2 are derived from different sources they appear to map very similar results. This lends credibility to their assessment. It suggests we could use either one safely or a combination of the two. One could weigh the third formula more heavily in a combination of the three if we believed that oaks do not add much biomass with additional diameter when they become very large. 
oakAGB.ave = function(b0,b1,d) exp(b0 + b1*log(d))

```

#Crown allometry
```{r}

```



#Choet et all on xylem
```{r}
choet <- read_csv('~/Desktop/Groundwater Research/Choat_et_al_2012_data.xls .csv') 
View(choet)
π#Lets look at the 50 loss by various groups in the data set that may be useful. 
names(choet)
unique(choet$Group)
unique(choet$Species)
#Get rid of the empty row
choet1 <- choet[2:nrow(choet),]
#lets fix these names
colnames(choet1)[3:8] <- c("xylem50","xylem88","xylem_min_mday", "xylem_min", "xylem50_safety", "xylem88_safety" )
choet1$Genus <- NA
for(i in 1:nrow(choet1)){
  choet1$Genus[i] <- strsplit(choet1$Species, " ")[[i]][1]
}
choet1$xylem50 <- as.numeric(choet1$xylem50)
choet1 %>% group_by(Genus) %>% summarise(mean(xylem50))
choet1 %>% group_by(Group) %>% summarise(mean(xylem50))



```

#Sequoia
```{r}
sequoia <- read_csv("~/Desktop/Groundwater Research/NPP_REDWOOD_803/data/Redwood_NPP.csv", 
    skip = 16)
#this needs a tremendous amount of cleaning and manipulation to become something useful. But it still may be in the future. 

```

Book called ecological forecasting book by Mike Ditze

Next step is to set the grids on the same resolution from the file in the google drive. Use Scypi to parse the netCDF files and get them to the same grid size. 
